Diagnostic Plots

### 2024-07-29 pre_print_manuscript.qmd
Additional study sims

https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(15)00489-4/fulltext

https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(15)01354-9/fulltext

@@

# Introduction
## Cumulative evidence synthesis 
Two questions that should be asked (though arguably are not often enough particularly in sport science) by researchers when planning a study of an experimental intervention is "what is the likelihood that the experimental intervention is superior to the control intervention given the evidence accumulated so far?", and "what is the likelihood that a new trial, given some design parameters and previous evidence, will demonstrate the superiority of the experimental intervention?".
The key here is to consider the *cumulative* nature of evidence provided by research and its synthesis. Indeed, to not do this could lead to so called "research waste". Evidence synthesis methods are essential to determining whether or not there is justification for further research on a given topic, and the Cochrane-Collaboration and REWARD (Reduce Research Waste and Reward Diligence) Alliance have even established an award for efforts in the area of reducing "research waste" [@glasziouResearchWasteStill2018].

          Add about QRPs and research waste - evaluate the not only the results, but the quality of the preceding evidence in addition to whether it may have been subject to any bias

Cumulative meta-analyses were proposed in the early 1990s and have since then been promoted as key tools to understand whether or not additional research is a worthwhile use of resources for addressing a particular question regarding experimental intervention [@clarkeAccumulatingResearchSystematic2014; @graingerEvidenceSynthesisTackling2020]. Further, Bayesian approaches are well positioned to tackle this [@biauUsingBayesianStatistics2017].

Within Bayesian statistical inference a prior probability distribution regarding the effect of interest is *updated* after the introduction of new evidence to a posterior probability distribution given Bayes theorem. In the present paper we demonstrate the application of Bayesian methods to meta-analysis in the sport and exercise sciences for the evaluation of experimental interventions using cumulative evidence synthesis. We assume some prior knowledge of evidence synthesis and meta-analytic methods on the part of the reader, though for those unfamiliar suggest some recent introductory papers regarding their application in the field [@gunnellQuestionsAnswersConducting2020; @haggerMetaanalysis2022; @steeleMetaanalysisVariationSport2023]. We use the example of self-talk interventions and their effects upon motor/sport performance given a quantitative evidence synthesis has not been conducted on this topic to the best of our knowledge since 2011 when Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] conducted their systematic review and meta-analysis.

## Self-talk interventions
Sport psychology as a broad field has focused on the theorising of psychological constructs that might impact upon performance, and the subsequent experimental testing of theoretically informed interventions to address these constructs.
For example, a recent umbrella review identified thirty meta-analyses exploring the effects of different sport psychology constructs upon performance, thirteen of them examining the effects of interventions, finding an overall standardised mean difference (SMD) for positive constructs of 0.51 \[95% confidence interval: 0.42, 0.58\] [@lochbaumSportPsychologyPerformance2022]. One construct with a long history of philosophical, theoretical, and empirical work [@geurtsMakingSenseSelf2018; @brinthauptSelftalkResearchChallenges2023; @latinjakSelfTalkInterdisciplinaryReview2023] that has been the target of considerable investigation in this field has been self-talk.

@@

Whilst self-talk as a field may have matured in the post-2011 years with theoretical advancements in conceptualisation of the construct and mediators of its effects on performance, efforts to improve operationalisation/measurement, and efforts to improve methodology used in studying self-talk [@brinthauptSelftalkResearchChallenges2023; @geurtsMakingSenseSelf2018; @hardyReflectionsMaturingResearch2018; @latinjakSelfTalkInterdisciplinaryReview2023; @vanraalteSelftalkReviewSportspecific2016; @latinjakSpeakingClearly102019], it could be argued that understanding of the effectiveness of self-talk *interventions* (referred to in modern literature as 'strategic' self-talk; @latinjakSpeakingClearly102019) was "mature" prior to 2011.
The effect estimate from the meta-analysis of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] might be considered by some to have been already fairly precise only spanning 0.2 SMD, and indeed aswell for many of the moderator estimates.

Many additional studies on self-talk interventions have been conducted since 2011. It may well be that such recent work has improved our estimates of the effects of self-talk interventions and what moderates their effectiveness, or indeed contributed to other areas of understanding of the construct of self-talk. But, it is a reasonable question to ask, given the limited time and resource for conducting research in the field of sport science and what we already know regarding these interventions, whether and to what extent these studies have advanced our understanding of their effects, or whether they have largely contributed to so called "research waste" [@graingerEvidenceSynthesisTackling2020; @glasziouResearchWasteStill2018].

## Aim of the present work
The aim of this present work is to demonstrate the application of Bayesian methods in evidence synthesis, in particular to meta-analysis, in the sport and exercise sciences for the evaluation of experimental interventions. Given that there has not been, to the best of our knowledge, a meta-analytic synthesis of the effects of self-talk interventions effects upon sport/motor performance since Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] it represents a prime topic to utilise as an example for these methods. Therefore, our aim was to conduct an updated systematic review and Bayesian meta-analysis replicating the search, inclusion, and models of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] in order to demonstrate the application of cumulative evidence synthesis methods including; consideration of the initial probability that a new study of the effects of self-talk interventions would shift our prior belief in their effectiveness, the application of priors taken from the previous meta-analysis to be updated by new studies identified to a new posterior estimate of effect, and consideration of other sources of research waste from questionable research practices such as possible publication bias.

@@

Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] did not explicitly state a process or strategy to formulating their research question and search methods.
However, we assumed that the PICO (Participants, Intervention, Comparator and Outcome) framework was implicitly used and, with that assumption, we adopted the following inclusion criteria based on their description.
Participants were healthy and of any performance level.
The intervention was instruction to engage in positive self-talk[^1].
The comparator was no self-talk or unrelated self-talk.
Outcomes were sport or motor task performance.
We included both between and within group experimental designs with either pre-post, or post-only measurements of performance as well as within group pre-post trials.

@@

Studies were obtained through electronic journal searches and review articles along with personal records and communication.
The following databases -- Sport Discus, PsycINFO, PsycARTICLES and Medline -- were selected through the EBSCO database to search for the keywords.
The SCOPUS database, used by Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011], was not used as it was not accessible through Solent University[^2].
These keywords were searched in the format of, with the application of the Boolean commands, (self-talk OR self-instruction OR self-statements OR self-verbalizations OR verbal cues OR stimulus cueing OR thought content instructions) AND (sport OR performance OR motor performance OR task performance).
The studies were all peer-reviewed, full text and published in English language journals.
The search was limited to the date range of November 2011-November 2023.
An initial search took place from October 2022-November 2022 as this project was completed as part of the lead authors undergraduate thesis.
We subsequently updated the search from November 2022-November 2023 prior to preparing this manuscript for publication.

[^2]: Though we feel fairly confident, given the number of studies identified and the findings of our models reported below, that any missed studies would be unlikely to qualitatively impact the overall findings and conclusions of this work. We should also note that we originally intended to report a PRISMA flow diagram of our search and retrieval process. However, for transparency, we encountered an issue where we realised that some of the searches we had conducted were not correctly recorded by the tool being used to manage the process. We were, despite attempting to do so, unable to exactly replicate the searches (number of hits from initial search string in databases used) at the time we realised from attempting to reproduce the searches again. As such, we do not present a PRISMA diagram though as noted we do feel fairly confident we have not missed any key studies nor that minor omissions would affect the overall results of our analyses.

@@

ADD DETAILS ABOUT QUALITATIVE APPRAISAL OF STUDY AIMS

@@

Statistical analysis of the data extracted was performed in R, (v 4.2.2; R Core Team, https://www.r-project.org/) and RStudio (v 2023.06.1; Posit, https://posit.co/).
As noted, we adopted a Bayesian approach to the present meta-analysis.
Specifically, we adopted informative priors from the models reported by Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] and replicated their analyses using appropriate model specifications given the structure of the data.
All code utilised for data preparation and analyses are available in either the Open Science Framework page for this project [https://osf.io/dqwh5/](https://osf.io/dqwh5/) or the corresponding GitHub repository [https://github.com/jamessteeleii/self_talk_meta_analysis_update](https://github.com/jamessteeleii/self_talk_meta_analysis_update).
The `renv` package [@usheyRenvProjectEnvironments2023] was used for package version reproducibility and a function based analysis pipeline using the `targets` package [@landauTargetsDynamicFunctionOriented2023] was employed (the analysis pipeline can be viewed by downloading the R Project and running the function `targets::tar_visnetwork()`). Standardised effect sizes were all calculated using the `metafor` package [@viechtbauerMetaforMetaAnalysisPackage2023]. The main package `brms` [@burknerBrmsBayesianRegression2023] was used in fitting all the Bayesian meta-analysis models. Prior and posterior draws were taken using `tidybayes` [@kayTidybayesTidyData2023] and `marginaleffects` [@arel-bundockMarginaleffectsPredictionsComparisons2023] packages. Bayes factors were calculated using the `bayestestR` package [@makowskiBayestestRUnderstandDescribe2023]. All visualisations were created using `ggplot2` [@wickhamGgplot2CreateElegant2023], `tidybayes`, and the `patchwork` [@pedersenPatchworkComposerPlots2023] packages.

Effect sizes were calculated as SMDs dependent on the design of the study. Firstly, all were signed such that a positive effect indicated that the self-talk intervention was favoured. For studies utilising a pretest-posttest-control comparison design we calculated the SMD between groups/conditions using the pooled pre-test standard deviation as per Morris [@morrisEstimatingEffectSizes2008]. For post-test only control comparison designs we calculated the SMD between groups/conditions based upon the pooled post-test standard deviation. Lastly, for single arm within group pre-post (or control-intervention) designs we calculated the SMD from pre- to post-intervention using the pre-test standard deviation. Effect sizes were all calculated using the `escalc` function from the `metafor` package.

Though it was not entirely clear from the reporting in the meta-analysis of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011], they noted including a greater number of effect sizes than individual studies. As such, it was likely that their data had a hierarchical structure with effects nested within studies whether they explicitly applied a hierarchical model to it or not. The studies we identified and included also had hierarchical structure whereby we had effects nested within groups (for example when there were multiple self-talk interventions examined) nested within experiments (for example when a study reported on multiple experiments using different samples and/or designs) nested within studies. As such, we used multilevel mixed effects meta-analyses with nested random intercepts for effects, groups, experiments, and studies. Effects were all weighted by the inverse sampling variance. A main model was produced which included all effects and was intended to update the overall model from Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011], whereby their overall estimate reflected the fixed model intercept. In addition, we produced models for each of the aforementioned categorical moderators where we excluded the model intercept in order to set priors for each category directly based on the estimates and their precision reported[^6] by Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011].

[^6]: It is not clear exactly from the reporting by Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] as to whether their moderator analyses involved fitting their models to datasets constructed separately, whereby they were filtered to only include the relevant category for the moderator, or if they were fit together including all categories. We opted to fit them together for parsimony here.

Priors for each model were set informatively for the intervention effects, and were set to be weakly regularising for the heterogeneity (i.e., $\tau$) at all levels of the model[^7]. Intervention effects were set with priors based on the effect estimates from Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011], reported in their Table 1, using a $t$-distribution ($t(k,\mu,\sigma)$) with $k-2$ degrees of freedom [@higginsReevaluationRandomeffectsMetaanalysis2009]. We assumed $k$ to be the number of effects included in the models reported by Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011]. For the main model the prior for the intervention effect was set directly on the model intercept i.e., $t(60, 0.48, 0.05)$. For the moderator models, as noted, we removed the model intercept allowing us to set the priors directly on each category for each moderator based on the estimates from Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] Table 1. In cases where moderators had new categories introduced in the newer studies, included in our analyses, we used $\mu=0.48$ and $\sigma=0.05$ taken from the overall estimate of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] and applied degrees of freedom $k=3$ to be more conservative and allow greater mass in the tails of the prior distribution for these categories. In all models the heterogeneity priors at each level were set using the default weakly regularising prior in `brms`; a half-$t$-distribution with $\mu=0$, $\sigma=2.5$, and $k=3$. This constrained the prior to only allow positively signed values for $\tau$ though over a wide range of possible values.

[^7]: Though Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] report on $\tau$ it is not clear what level this applies to (and as noted it is not clear if they employed a hierachical model) and they do not report any interval estimate for this making it difficult to specify an informative prior distribution. As such, and given suggestions regarding heterogeneity priors [@williamsBayesianMetaAnalysisWeakly2018; @roverWeaklyInformativePrior2021], we opted for a weakly regularising distribution at all levels.

As we were interested in determining how much the new evidence produced since Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] had updated our understanding of the effects of self-talk interventions, we fit each model using four Monte Carlo Markov Chains each with 4000 warmup and 40000 sampling iterations. This was in order to obtain precise Bayes Factors using the Savage-Dickey ratio [@gronauBridgesamplingPackageEstimating2020]. Trace plots were produced along with $\hat{R}$ values to examine whether chains had converged, and posterior predictive checks for each model were also examined to understand the model implied distributions. These all showed good convergence with all $\hat{R}$ values close to 1 and  posterior predictive checks seemed appropriate distributions for the observed data (all diagnostic plots can be seen in the supplementary materials: [https://osf.io/xbcw8](https://osf.io/xbcw8)).

From each model we obtained draws from the posterior distributions for the intervention effects (i.e., the expectation of the value of the parameters posterior probability distribution) in order to present probability density functions visually, and also to calculate mean and 95% quantile intervals (i.e., 'credible' or 'compatibility' intervals) for each estimate. These gave us the most probable value of the parameter in addition to the range from 2.5% to 97.5% percentiles. The same was done drawing samples from the prior distributions only in order to present both distributions visually for comparison of the prior to posterior updating. For the main model draws were taken at the study level and an ordered forest plot produced showing each studies posterior distribution along with mean and 95% quantile intervals. We also calculated the 95% prediction intervals providing the range over which we can expect 95% of future effect estimates to fall and present each individual effect size on the forest plot. To compliment the visual inspection of prior to posterior updating we also present log10 Bayes Factors (log10[BF]) calculated *against* 100 effects ranging from a SMD of 0 through to 1 and plot these log10(BF) curves for each model intervention effect estimate i.e.., the Savage-Dickey ratio was calculated for each of 100 point effects in the interval (0,1) equally spaced. These were compared to Jeffreys [-@jeffreysTheoryProbability1998] scale regarding evidence *against* (i.e., 0 to 0.5 = weak evidence; 0.5 to 1 = substantial evidence; 1 to 1.5 = strong evidence; 1.5 to 2 = very strong evidence; 2 or greater = decisive evidence). Thus, a positive log10(BF) value indicated that, compared to the prior distribution (meaning the estimates of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011]), there was now greater evidence against the SMD for which the log10(BF) was calculated. A loess smooth was then applied to these 100 values.

>>>>

@@

# Introduction
## Cumulative evidence synthesis 
Two questions that should be asked (though arguably are not asked often enough particularly in sport science) by researchers when planning a study of an experimental intervention is "what is the likelihood that the experimental intervention is superior to the control intervention given the evidence accumulated so far?", and "what is the likelihood that a new trial, given some design parameters and previous evidence, will demonstrate the superiority of the experimental intervention?". The key here is to consider the *cumulative* nature of evidence provided by research and its synthesis. Indeed, to not do this could lead to redundancy or so called "research waste". Evidence synthesis methods are essential to determining whether or not there is justification for further research on a given topic, and the Cochrane-Collaboration and REWARD (Reduce Research Waste and Reward Diligence) Alliance have even established an award for efforts in the area of reducing "research waste" [@glasziouResearchWasteStill2018]. However, across many domains there remains a high prevalence of redundancy and a low prevalence of attempts to minimise or reduce it [@lundMetaresearchEvaluatingRedundancy2022].

Cumulative meta-analyses were proposed in the early 1990s and have since then been promoted as key tools to understand whether or not additional research is a worthwhile use of resources for addressing a particular question regarding experimental intervention [@clarkeAccumulatingResearchSystematic2014; @graingerEvidenceSynthesisTackling2020]. Further, Bayesian approaches are well positioned to tackle this [@biauUsingBayesianStatistics2017]. Within Bayesian statistical inference a prior probability distribution regarding the effect of interest is *updated* after the introduction of new evidence to a posterior probability distribution given Bayes theorem.

The trustworthiness of prior data should also be considered in evidence synthesis. Meta-analyses rely on the assumption that the sample of studies included is not based on a biased selection procedure either on the part of the systematic reviewer(s) or with regards to the studies present in the literature to sample from. For the latter, publication bias and *p*-hacking are the two most common phenomena that violate this assumption and can substantial influence cumulative evidence [@friesePHackingPublicationBias2020]. Publication bias is typically explored using methods such as selection models based on significance thresholds for *p*-values [@mcshaneAdjustingPublicationBias2016], funnel plot based regression methods [@stanleyMetaregressionApproximationsReduce2014], or methods which combine these approaches and leverage the uncertainty in the underlying true data generating process such as Robust Bayesian Meta-Analysis with model averaging [@bartosAdjustingPublicationBias2022; @bartosRobustBayesianMetaanalysis2023]. For *p*-hacking mixture models have recently been proposed [@mossModellingPublicationBias2023]. The existence of evidence suggesting prior questionable research practices such as publication bias or *p*-hacking in previous literature might be cause for an evaluation of whether a research program regarding an experimental intervention is worth building upon, or starting afresh incorporating safe-guards for such issues e.g., pre-registration/registered-reports [@chambersPresentFutureRegistered2022].

In the present paper we demonstrate the application of methods for cumulative evidence synthesis including Bayesian meta-analysis, and exploration of questionable research practices such as publication bias or *p*-hacking, in the sport and exercise sciences for the evaluation of experimental interventions. We assume some prior knowledge of evidence synthesis and meta-analytic methods on the part of the reader, though for those unfamiliar suggest some recent introductory papers regarding their application in the field [@gunnellQuestionsAnswersConducting2020; @haggerMetaanalysis2022; @steeleMetaanalysisVariationSport2023]. We use the example of self-talk interventions and their effects upon motor/sport performance given a quantitative evidence synthesis has not been conducted on this topic to the best of our knowledge since 2011 when Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] conducted their systematic review and meta-analysis, and thus is ripe to use in demonstrating cumulative methods such as Bayesian updating. Note, we do not intend to present this paper as a comprehensive systematic review and meta-analysis of self-talk interventions in sport/motor performance nor a thorough theoretical review of the construct. We do however present a brief overview of the topic below for context of this as an example.

## Self-talk interventions
Sport psychology as a broad field has focused on the theorising of psychological constructs that might impact upon performance, and the subsequent experimental testing of theoretically informed interventions to address these constructs and subsequent performance.
For example, a recent umbrella review identified thirty meta-analyses exploring the effects of different sport psychology constructs upon performance, thirteen of them examining the effects of interventions, finding an overall standardised mean difference (SMD) for positive constructs of 0.51 \[95% confidence interval: 0.42, 0.58\] [@lochbaumSportPsychologyPerformance2022]. One construct with a long history of philosophical, theoretical, and empirical work [@geurtsMakingSenseSelf2018; @brinthauptSelftalkResearchChallenges2023; @latinjakSelfTalkInterdisciplinaryReview2023] that has been the target of considerable investigation in this field has been self-talk.

@@

Whilst self-talk as a field may have matured in the post-2011 years with theoretical advancements in conceptualisation of the construct and mediators of its effects on performance, efforts to improve operationalisation/measurement, and efforts to improve methodology used in studying self-talk [@brinthauptSelftalkResearchChallenges2023; @geurtsMakingSenseSelf2018; @hardyReflectionsMaturingResearch2018; @latinjakSelfTalkInterdisciplinaryReview2023; @vanraalteSelftalkReviewSportspecific2016; @latinjakSpeakingClearly102019], it could be argued that understanding of the effectiveness of self-talk *interventions* (referred to in modern literature as 'strategic' self-talk; @latinjakSpeakingClearly102019) was "mature" prior to 2011. The effect estimate from the meta-analysis of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] might be considered by some to have been already fairly precise only spanning 0.2 SMD, and indeed aswell for many of the moderator estimates. Indeed, some authors had moved onto to attempting to explain *why* self-talk interventions are effective, exploring possible mechanisms with the starting assumption that these interventions had been proven as effective for enhancing performance [@galanisChapterWhySelfTalk2016].

Despite this, many additional studies on self-talk interventions have been conducted since 2011. It may well be that such recent work has further improved our estimates of the effects of self-talk interventions and what moderates their effectiveness, or indeed contributed to other areas of understanding of the construct of self-talk. But, it is a reasonable question to ask, given the limited time and resource for conducting research in the field of sport science and what we might claim to have already known regarding these interventions, whether and to what extent these studies have advanced our understanding of their effects, or whether they have largely contributed to so called "research waste" [@graingerEvidenceSynthesisTackling2020; @glasziouResearchWasteStill2018].

## Aim of the present work
The aim of this present work is to demonstrate the application of methods for cumulative evidence synthesis including Bayesian meta-analysis, and exploration of questionable research practices such as publication bias or *p*-hacking, in the sport and exercise sciences for the evaluation of experimental interventions. Given that there has not been, to the best of our knowledge, a meta-analytic synthesis of the effects of self-talk interventions effects upon sport/motor performance since Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] it represents a prime topic to utilise as an example for these methods. Therefore, our aim was to conduct an updated systematic review and Bayesian meta-analysis replicating the search, inclusion, and models of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] in order to demonstrate the application of cumulative evidence synthesis methods including; consideration of the initial probability that a new study of the effects of self-talk interventions would shift our prior belief in their effectiveness, the application of priors taken from the previous meta-analysis to be updated by new studies identified to a new posterior estimate of effect, and consideration of other sources of research waste from questionable research practices such as possible publication bias and *p*-hacking.

@@

Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] did not explicitly state a process or strategy to formulating their research question and search methods. However, we assumed that the PICO (Participants, Intervention, Comparator and Outcome) framework was implicitly used and, with that assumption, we adopted the following inclusion criteria based on their description. Participants were healthy and of any performance level. The intervention was instruction to engage in positive self-talk[^1]. The comparator was no self-talk or unrelated self-talk.
Outcomes were sport or motor task performance. We included both between and within group experimental designs with either pre-post, or post-only measurements of performance as well as within group pre-post trials similarly to Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011].

@@

Studies were obtained through electronic journal searches and review articles along with personal records and communication.
The following databases -- Sport Discus, PsycINFO, PsycARTICLES and Medline -- were selected through the EBSCO database to search for the keywords.
The SCOPUS database, used by Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011], was not used as it was not accessible through Solent University[^2]. These keywords were searched in the format of, with the application of the Boolean commands, (self-talk OR self-instruction OR self-statements OR self-verbalizations OR verbal cues OR stimulus cueing OR thought content instructions) AND (sport OR performance OR motor performance OR task performance). The studies were all peer-reviewed, full text and published in English language journals.
The search was limited to the date range of November 2011-November 2023. An initial search took place from October 2022-November 2022 as this project was completed as part of the lead authors undergraduate thesis. We subsequently updated the search from November 2022-November 2023 prior to initially preparing this manuscript for publication.

[^2]: Though we feel fairly confident, given the number of studies identified and the findings of our models reported below, that any missed studies would be unlikely to qualitatively impact the overall findings and conclusions of this work. We should also note that we originally intended to report a PRISMA flow diagram of our search and retrieval process. However, for transparency, we encountered an issue where we realised that some of the searches we had conducted were not correctly recorded by the tool being used to manage the process. We were, despite attempting to do so after realising this, unable to exactly replicate the searches (number of hits from initial search string in databases used) at the time we realised from attempting to reproduce the searches again. As such, we do not present a PRISMA diagram though as noted we do feel fairly confident we have not missed any key studies nor that minor omissions would affect the overall results of our analyses anyway. Further, as noted we have chosen this as an example to demonstrate the methods moreso than to conduct a comprehensive systematic review of the topic.

@@

@@

Statistical analysis of the data extracted was performed in R, (v 4.2.2; R Core Team, https://www.r-project.org/) and RStudio (v 2023.06.1; Posit, https://posit.co/). All code utilised for data preparation and analyses are available in either the Open Science Framework page for this project [https://osf.io/dqwh5/](https://osf.io/dqwh5/) or the corresponding GitHub repository [https://github.com/jamessteeleii/self_talk_meta_analysis_update](https://github.com/jamessteeleii/self_talk_meta_analysis_update).

ADD grateful citation report

    The `renv` package [@usheyRenvProjectEnvironments2023] was used for package version reproducibility and a function based analysis pipeline using the `targets` package [@landauTargetsDynamicFunctionOriented2023] was employed (the analysis pipeline can be viewed by downloading the R Project and running the function `targets::tar_visnetwork()`). Standardised effect sizes were all calculated using the `metafor` package [@viechtbauerMetaforMetaAnalysisPackage2023]. The main package `brms` [@burknerBrmsBayesianRegression2023] was used in fitting all the Bayesian meta-analysis models. Prior and posterior draws were taken using `tidybayes` [@kayTidybayesTidyData2023] and `marginaleffects` [@arel-bundockMarginaleffectsPredictionsComparisons2023] packages. Bayes factors were calculated using the `bayestestR` package [@makowskiBayestestRUnderstandDescribe2023]. All visualisations were created using `ggplot2` [@wickhamGgplot2CreateElegant2023], `tidybayes`, and the `patchwork` [@pedersenPatchworkComposerPlots2023] packages.

### Examining the effects of a new trial upon belief in the effects of self-talk interventions
To begin with we examined through simulation what impact a single new trial might have had upon shifting belief in the prior estimate yielded by the meta-analysis of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011]. Sample size was simulated as 10, 20, 40, 80, 160, 320, 640, 1282, 2560, and 5120[^6] with a 50:50 allocation to either self-talk intervention or control conditions, and we varied the sample effect size as an SMD of 0, 0.2, 0.4, 0.6, 0.8, and 1.0 reflecting a range from no effect of self-talk interventions to a large  effect. In each combination of sample size and true effect size we set the sample parameters to  the SMD and its corresponding sampling variance was calculated and then included as a single observation in a Bayesian random effects meta-analysis where the priors was set informatively for the intervention effects, and were set to be default weakly regularising for the heterogeneity (i.e., $\tau$)[^7]. Intervention effects were set with priors based on the effect estimates from Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011], reported in their Table 1, using a $t$-distribution ($t(k,\mu,\sigma)$) with $k-2$ degrees of freedom [@higginsReevaluationRandomeffectsMetaanalysis2009]. We assumed $k$ to be the number of effects included in the models reported by Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011]. The prior for the intervention effect was set directly on the model intercept i.e., $t(60, 0.48, 0.05)$. The prior for heterogeneity was set as the default weakly regularising prior in `brms`; a half-$t$-distribution with $\mu=0$, $\sigma=2.5$, and $k=3$. This constrained the prior to only allow positively signed values for $\tau$ though over a wide range of possible values. We fit each model using four Monte Carlo Markov Chains each with 2000 warmup and 6000 sampling iterations.

From each model we obtained draws from the posterior distributions for the intervention effects (i.e., the expectation of the value of the parameters posterior probability distribution) in order to present probability density functions visually. The same was done drawing samples from the prior distributions only in order to present both distributions visually for comparison of the prior to posterior updating. As a means of examining the extent to which the posterior distribution for the self-talk intervention effect estimate was shifted from the prior distribution as a result of introducing each new trial we calculated the proportion of the full posterior distribution within the 95% quantile interval of the prior distribution i.e., the most probable value of the parameter in addition to the range from 2.5% to 97.5% percentiles (equivalent to the 95% confidence interval of the estimate from Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011]), using the a Region of Practical Equivalence approach [@kruschkeBayesianNewStatistics2018a]. In essence, where the proportion of the posterior distribution that was within the 95% quantile interval from the prior distribution was ~95% then we would conclude that the new trial had little impact on shifting our prior belief in the intervention effect. This helps understand whether it might be worthwhile to conduct the kind of study that would need to be conducted in terms of sample size, and assuming what the true effect might be, or whether doing so might be a waste of resources given the precision of existing estimates of the intervention effect.

[^6]: Sample sizes were doubled up to roughly a similar sample size as the largest study post-2011 that we identified in our updated searches [@laneBriefOnlineTraining2016].

[^7]: Though Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] report on $\tau$ it is not clear what level this applies to (and as noted it is not clear if they employed a hierachical model) and they do not report any interval estimate for this making it difficult to specify an informative prior distribution. As such, and given suggestions regarding heterogeneity priors [@williamsBayesianMetaAnalysisWeakly2018; @roverWeaklyInformativePrior2021], we opted for a weakly regularising distribution at all levels including the updated multilevel meta-analysis described further in these methods and in the random effects meta-analysis used in these simulations.

### Updating the prior esitmate from Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] with newer studies
For all the groups/conditions in studies identified in our updated searches effect sizes were calculated as SMDs dependent on the design of the study. Firstly, all were signed such that a positive effect indicated that the self-talk intervention was favoured. For studies utilising a pretest-posttest-control comparison design we calculated the SMD between groups/conditions using the pooled pre-test standard deviation as per Morris [-@morrisEstimatingEffectSizes2008]. For post-test only control comparison designs we calculated the SMD between groups/conditions based upon the pooled post-test standard deviation. Lastly, for single arm within group pre-post (or control-intervention) designs we calculated the SMD from pre- to post-intervention using the pre-test standard deviation.

Though it was not entirely clear from the reporting in the meta-analysis of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011], they noted including a greater number of effect sizes than individual studies. As such, it was likely that their data had a hierarchical structure with effects nested within studies whether they explicitly applied a hierarchical model to it or not. The studies we identified and included also had hierarchical structure whereby we had effects nested within groups (for example when there were multiple self-talk interventions examined) nested within experiments (for example when a study reported on multiple experiments using different samples and/or designs) nested within studies. As such, we used multilevel mixed effects meta-analyses with nested random intercepts for effects, groups, experiments, and studies. Effects were all weighted by the inverse sampling variance. A main model was produced which included all effects and was intended to update the overall model from Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011], whereby their overall estimate reflected the fixed model intercept. In addition, we produced models for each of the aforementioned categorical moderators where we excluded the model intercept in order to set priors for each category directly based on the estimates and their precision reported (see footnote$^7$) by Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011].

Priors for each model were again set informatively for the intervention effects, and were set to be weakly regularising for the heterogeneity (i.e., $\tau$) at all levels of the model. Intervention effects were set with priors based on the effect estimates from Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011], reported in their Table 1, using a $t$-distribution ($t(k,\mu,\sigma)$) with $k-2$ degrees of freedom [@higginsReevaluationRandomeffectsMetaanalysis2009]. We assumed $k$ to be the number of effects included in the models reported by Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011]. For the main model the prior for the intervention effect was set directly on the model intercept i.e., $t(60, 0.48, 0.05)$. For the moderator models, as noted, we removed the model intercept allowing us to set the priors directly on each category for each moderator based on the estimates from Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] Table 1. In cases where moderators had new categories introduced in the newer studies, included in our analyses, we used $\mu=0.48$ and $\sigma=0.05$ taken from the overall estimate of Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] and applied degrees of freedom $k=3$ to be more conservative and allow greater mass in the tails of the prior distribution for these categories. In all models the heterogeneity priors at each level were set using the default weakly regularising prior in `brms`; a half-$t$-distribution with $\mu=0$, $\sigma=2.5$, and $k=3$. This constrained the prior to only allow positively signed values for $\tau$ though over a wide range of possible values.

As we were interested in determining how much the new evidence produced since Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] had updated our belief in the effects of self-talk interventions, we fit each model using four Monte Carlo Markov Chains each with 4000 warmup and 40000 sampling iterations. This was in order to obtain precise Bayes Factors using the Savage-Dickey ratio [@gronauBridgesamplingPackageEstimating2020]. Trace plots were produced along with $\hat{R}$ values to examine whether chains had converged, and posterior predictive checks for each model were also examined to understand the model implied distributions. These all showed good convergence with all $\hat{R}$ values close to 1 and  posterior predictive checks seemed appropriate distributions for the observed data (all diagnostic plots can be seen in the supplementary materials: [https://osf.io/xbcw8](https://osf.io/xbcw8)).

From each model we obtained draws from the posterior distributions for the intervention effects (i.e., the expectation of the value of the parameters posterior probability distribution) in order to present probability density functions visually, and also to calculate mean and 95% quantile intervals (i.e., 'credible' or 'compatibility' intervals) for each estimate. These gave us the most probable value of the parameter in addition to the range from 2.5% to 97.5% percentiles. The same was done drawing samples from the prior distributions only in order to present both distributions visually for comparison of the prior to posterior updating. For the main model draws were taken at the study level and an ordered forest plot produced showing each studies posterior distribution along with mean and 95% quantile intervals. We also calculated the 95% prediction intervals providing the range over which we can expect 95% of future effect estimates to fall and present each individual effect size on the forest plot.
